---
title: "balanced_social_behavior_data_manipulation"
author: "Yama Chang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: no
      smooth_scroll: no
geometry: margin=0.50in
---

# ============================================================================================= #
# ============================================================================================= #
# Tracking Depression Project
# Balancing Real-Life and Virtual Interactions: How Social Behavior Predicts Depressive Symptoms
# ============================================================================================= #
# ============================================================================================= #


# Set up

```{r setup, include=FALSE, cache = FALSE}
require("knitr")
getwd() # Check where's your working directory
## setting working directory if not using git
## setwd("") # Remove hashtag at the beginning and add your working directory inside ""
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, warning = FALSE, message = FALSE, include = TRUE, cache = FALSE)
```

```{r load packages, include=FALSE}

if(!require(car)){install.packages('car')}
library(car)
if(!require(psych)){install.packages('psych')}
library(psych)
if(!require(effects)){install.packages('effects')}
library(effects)
if(!require(tidyverse)){install.packages('tidyverse')}
library(tidyverse)
if(!require(summarytools)){install.packages('summarytools')}
library(summarytools)
if(!require(diffdf)){install.packages('diffdf')}
library(diffdf)
if(!require(janitor)){install.packages('janitor')}
library(janitor)
if(!require(haven)){install.packages('haven')}
library(haven)
if(!require(lavaan)){install.packages('lavaan')}
library(lavaan)
if(!require(readxl)){install.packages('readxl')}
library(readxl)
if(!require(lubridate)){install.packages('lubridate')}
library(lubridate)
if(!require(mice)){install.packages('mice')}
library(mice)
if(!require(here)){install.packages('here')}
library(here)

```

```{r parameters}
# ---- Parameters ----
STUDY_DAYS <- 90
IMPUTATION_M <- 5
IMPUTATION_SEED <- 333
SOCIAL_APP_TIMEOUT_SEC <- 5 * 60  # 5-minute timeout for app duration calculation
APP_TIMEZONE <- "UTC"
```

# Data Cleaning

[V] Outcome variable: EMA data (PHQ9)
[V] Online social interaction
  - [V] calllog: `uid`, `date`, `total_duration`
  - [V] smslog: 
  - [V] appusage
    - [V] running_app_123
    - [V] unlocked
[V] In-Person social interaction
  - [V] audio_conversation (ALL pts): `uid`, `date`, `total_convo_length_sec`, `total_convo_length_min`
  - [X] audio_inference (ALL pts)

## EMA data
```{r}
# Load data with 300 TD participants
data_ema <- read_csv(here::here("data", "3._TD_cohort_EMA_FINAL.csv")) %>%
  janitor::clean_names() # convert column names to lower snake case

# Create a ID list for these 300 pts
IDlist <- unique(data_ema$mlife_id)

data_ema_single_check <- data_ema %>% filter(mlife_id == "REDACTED_ID_1@example")
```

I need to fix the total score of PHQ-9 by:

1) Missingness imputation before calculating the "total score".
2) Get the mean (OR total?) of the daily PHD-9 score.
3) Substitute the final PHQ-9 in the final df with this new column.
4) Rerun the analysis.

### Calculating daily PHQ-9
```{r}

# Select columns for this anlaysis
data_ema_1 <- data_ema %>% select(uid = mlife_id, date = date_rep, ema_time, ema_timestamp, matches("phq"))

# Transform date format
data_ema_1$date <- as.Date(data_ema_1$date, format = "%m/%d/%Y")

```

If the time span between the first and last date for a participant is 90 days or more, keep only those consecutive dates.

If the time span is less than 90 days, extend the sequence to exactly 90 days from the first available date.

```{r bin dates for EMA data}

# Step 1: Let's check what we're working with
check_dates <- data_ema_1 %>%
  group_by(uid) %>%
  summarise(
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE)
  )

# Print to see what we're working with
print(check_dates)

# Step 2: Create the complete sequence
data_ema_2 <- data_ema_1 %>%
  # First get a single row per participant with date range info
  group_by(uid) %>%
  summarise(
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE)
  ) %>%
  # Calculate timespan and determine end date
  mutate(
    time_span = as.numeric(difftime(last_date, first_date, units = "days")),
    end_date = if_else(
      time_span >= 90,
      last_date,
      first_date + lubridate::days(89)
    )
  ) %>%
  # Now create the sequence for each participant
  rowwise() %>%  
  mutate(
    date_seq = list(seq.Date(
      from = first_date,
      to = end_date,
      by = "day"
    ))
  ) %>%
  unnest(date_seq) %>%
  # Add bin_dates column that counts days from start (1-based)
  group_by(uid) %>%
  mutate(
    bin_dates = row_number()  # This will number days sequentially starting from 1
  ) %>%
  ungroup() %>%
  select(uid, date = date_seq, bin_dates, first_date, end_date) %>%
  # Join back with original data
  left_join(data_ema_1, by = c("uid", "date"))

# Step 1: Let's check what we're working with
check_dates <- data_ema_2 %>%
  group_by(uid) %>%
  summarise(
    min_date = min(bin_dates, na.rm = TRUE),
    max_date = max(bin_dates, na.rm = TRUE)
  )

# Print to see what we're working with
print(check_dates)
```


```{r ema data imputation}
# Perform multiple imputation
ema_method <- make.method(data_ema_2)
ema_method[] <- ""
ema_method[paste0("phq_", 1:9)] <- "pmm"

ema_pred <- make.predictorMatrix(data_ema_2)
ema_pred[,] <- 0
ema_pred[paste0("phq_", 1:9), c("bin_dates", paste0("phq_", 1:9))] <- 1
diag(ema_pred) <- 0

mice_ema <- mice(
  data_ema_2,
  m = IMPUTATION_M,
  method = ema_method,
  predictorMatrix = ema_pred,
  remove.collinear = FALSE,
  seed = IMPUTATION_SEED
) # For reproducibility

# Extract the first imputed dataset
data_ema_3 <- complete(mice_ema, 1)

# Keep the first 90 days
data_ema_4 <- data_ema_3 %>% filter(bin_dates %in% 1:STUDY_DAYS)

sum(is.na(data_ema_4$phq_1)) # check NA after imputation
```


```{r}
# Total PHQ-9
data_ema_5 <- data_ema_4 %>%
  mutate(phq_total = rowSums(select(., paste0("phq_", 1:9)), na.rm = TRUE),
         phq_mean = rowMeans(select(., paste0("phq_", 1:9)), na.rm = TRUE))

# Daily PHQ-9
data_ema_6 <- data_ema_5 %>%
  group_by(uid, date) %>%
  summarize(
    daily_phq_total = mean(phq_total, na.rm = TRUE),
    daily_phq_mean = mean(phq_mean, na.rm = TRUE),
    bin_dates = first(bin_dates),
    .groups = "drop"
  )

# Export as csv
```

## Sensing data 
```{r read data, include=FALSE}
# Note: complete_feat_dataset.csv and app_usage_feat.csv were earlier
# pre-processed feature files used for exploratory inspection. The pipeline
# below reads each data stream from its original source file (calllog.csv,
# df_convo_3_02202025.csv, df_running_app_combined.csv, etc.), so this
# chunk is no longer needed.
```

### Battery data

```{r}
# List all CSV files in the "convo" folder
file_paths <- list.files(path = here::here("data", "battery"), pattern = "*.csv", full.names = TRUE)

df_battery <- file_paths %>%
  map_dfr(~ read_csv(.x) %>%
            clean_names() %>%
            mutate(across(everything(), as.character))) 

df_battery_1 <- df_battery %>% 
  select(uid, day, no = x1, battery = data) %>% 
  mutate(
    date = as.Date(day),               # Extract the date (yyyy-mm-dd)
    hour = hour(ymd_hms(day))          # Extract the hour of the day
  ) %>%
  distinct(uid, date, hour, .keep_all = TRUE)  # Make sure unique rows per date and hour

# Create a complete grid for each `uid`, `date`, and `hour`
df_battery_complete <- df_battery_1 %>%
  # First get clean date ranges for each uid
  group_by(uid) %>%
  summarise(
    min_date = min(date, na.rm = TRUE),
    max_date = max(date, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Create nested dataframe with date sequences
  rowwise() %>%
  mutate(date = list(seq(min_date, max_date, by = "day"))) %>%
  unnest(date) %>%
  # Join back with original data
  left_join(df_battery_1, by = c("uid", "date")) %>%
  # Now complete the hours
  group_by(uid, date) %>%
  complete(
    hour = 0:23,
    fill = list(no = NA, battery = NA, day = NA)
  ) %>%
  select(-min_date, -max_date) %>%
  ungroup()

# df_battery_complete <- df_battery_1 %>%
#   group_by(uid, date) %>%
#   complete(
#     hour = 0:23, # Ensure all 24 hours are included
#     fill = list(no = NA, battery = NA) # Fill missing combinations with NA
#   ) %>%
#   ungroup()
test_battery1 <- df_battery %>% filter(uid == "REDACTED_ID_1@example")

df_battery_complete <- df_battery_complete %>%
  mutate(
    running = ifelse(is.na(battery), 0, 1) # Code as 0 for NA and 1 for non-NA
  )

# Aggregate data to day-level and create `running_day`
df_battery_day_level <- df_battery_complete %>%
  group_by(uid, date) %>%
  summarize(
    running_day = ifelse(any(running == 1, na.rm = TRUE), 1, 0), # Check if at least one `running == 1`
    .groups = "drop"
  )

# Check for missing dates and bin the data
df_battery_day_level_1 <- df_battery_day_level %>%
  group_by(uid) %>%
  # Create a complete sequence of dates for each participant
  complete(date = seq(min(date), max(date), by = "1 day")) %>%
  # Add a "bin" column for days starting from the earliest date
  mutate(bin = as.numeric(difftime(date, min(date), units = "days")) + 1) %>% 
  ungroup()

# Count NA values in the `running_day` column
sum(is.na(df_battery_day_level_1$running_day))


# Check for missing days in the sequence 1 to 90 for each participant
df_battery_day_level_1 %>%
  group_by(uid) %>%
  summarize(
    missing_days = list(setdiff(1:90, bin)), # Compare expected vs actual sequence
    total_missing = length(setdiff(1:90, bin)) # Count missing days
  ) %>%
  filter(total_missing > 0) # Keep only participants with missing days

```

```{r}
# check for duplicates in the Battery data per date & hour combination:
duplicates_detail <- df_battery_complete %>%
  group_by(uid, date, hour) %>%
  filter(n() > 1) %>%
  arrange(uid, date, hour)

duplicates_detail
```

```{r}
# Identify participants with missing hours
missing_hours_summary <- df_battery_complete %>%
  filter(is.na(battery)) %>%
  group_by(uid, date) %>%
  summarize(missing_hours = n(), .groups = "drop")

print(missing_hours_summary)

```




### Call Log

Total time of call duration.

Q: `type` data has more than 7 categories. Select only `1. Incoming`, `2. Outgoing`, and `7: Answered externally`

1: Incoming; 2: Outgoing; 3: Missed; 4: Voicemail; 5: Rejected; 6: Blocked; 7: Answered externally (Call log type for a call which was answered on another device. Used in situations where a call rings on multiple devices simultaneously and it ended up being answered on a device other than the current one.)

```{r}
df_calllog <- read_csv(here::here("data", "calllog.csv")) %>%
  janitor::clean_names() # convert column names to lower snake case

colnames(df_calllog)

# Select columns that'd be helpful for this analysis
df_calllog_1 <- df_calllog %>% select(uid, duration:type) 
```
```{r}
# Step 1: Get clean date ranges from data_ema_6 (one row per uid)
date_ranges <- data_ema_6 %>%
  group_by(uid) %>%
  summarise(first_date = min(date), end_date = max(date), .groups = "drop")

# Step 2: Process initial call log data
df_calllog_base <- df_calllog_1 %>% 
  mutate(
    date = as.Date(time),
    hour = hour(ymd_hms(time)),
    hour = ifelse(is.na(hour), 0, hour)
  )

# Step 3: Create template with all combinations and bin_dates
template <- date_ranges %>%
  rowwise() %>%
  mutate(
    dates = list(seq(first_date, end_date, by = "day"))
  ) %>%
  unnest(dates) %>%
  group_by(uid) %>%
  mutate(
    bin_dates = row_number()  # Add bin_dates
  ) %>%
  filter(bin_dates <= 90) %>%  # Keep only first 90 days
  crossing(hour = 0:23) %>%
  select(uid, date = dates, hour, bin_dates)

# Step 4: Join template with call data
df_calllog_1_complete <- template %>%
  left_join(
    df_calllog_base,
    by = c("uid", "date", "hour")
  ) %>%
  arrange(uid, date, hour)

# Verify the results
verification <- df_calllog_1_complete %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    n_hours = n() / n_distinct(date),
    .groups = "drop"
  )

print(head(verification))  # Should now show exactly 90 dates for all UIDs
```

```{r}
# `df_battery_complete` `df_calllog_1_complete`
# Check missing battery data and update call duration totals
df_calllog_1_battery <- df_calllog_1_complete %>%
  # Join with battery data to check if battery is missing
  left_join(df_battery_complete, by = c("uid", "date", "hour")) %>%
  # Create the `isbatterymissing` column: TRUE if no matching battery data
  mutate(battery_running = ifelse(is.na(battery), 0, 1)) %>%
  # Update `duration_new` to 0 if:
  # - Battery is running, AND
  # - `duration` is NA
  mutate(
    duration_new = ifelse(battery_running == 1 & is.na(duration), 0, duration)
                          #ifelse(battery_running == 0 & !is.na(duration), NA, duration))
  ) %>% 
  select(uid, date, hour, bin_dates, duration_new, time, type, battery_running)

# Verify the results
verification <- df_calllog_1_battery %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    n_hours = n() / n_distinct(date),
    .groups = "drop"
  )

# table(df_calllog_1_battery$duration_new)
# sum(is.na(df_calllog_1_battery$duration_new)) # original
# sum(is.na(df_calllog_1_battery$duration_new)) # new
# nrow(df_calllog_1_battery)

# Check the duration when battery's off: we might have call log data for longer than the study period while batter data is just the study period
df_calllog_1_batteryoff <- df_calllog_1_battery %>% filter(battery_running == 0)
table(df_calllog_1_batteryoff$duration_new)

# type data has weird value: it's less than 100 values - we can safely transform to NA
table(df_calllog_1_battery$type)

# Clean type data
df_calllog_1_battery_1 <- df_calllog_1_battery %>% 
  # Recode values other than 1 to 7 to NA
  mutate(type = ifelse(type %in% c(1, 2, 3, 4, 5, 6, 7), type, NA))

table(df_calllog_1_battery_1$type) # check!
```

#### Multiple imputation (Call log)
```{r Multiple imputation - call log}
# Fix: For missing dates with other types of calls: 0 secs - don’t impute
# sum(is.na(df_calllog_1_battery_1$duration_new)) check how may NAs

df_calllog_1_battery_2 <- df_calllog_1_battery_1 %>% 
  mutate(duration_new = ifelse(type %in% c(3, 4, 5, 6), 0, duration_new))

# maintain the complete date structure (all dates for each UID) even after filtering for specific types (1, 2, 7)
# First create a complete date frame from UNFILTERED data
complete_frame <- df_calllog_1_battery_2 %>%  # using original unfiltered data
  distinct(uid, date) %>%
  group_by(uid) %>%
  complete(date = seq(min(date), max(date), by = "1 day"))

# Then join with your filtered data
df_calllog_1_battery_3 <- df_calllog_1_battery_2 %>%
  filter(type %in% c(1, 2, 7)) %>%
  right_join(complete_frame, by = c("uid", "date")) %>% 
  arrange(uid, date)

df_filtered <- df_calllog_1_battery_3 %>%
  mutate(
    period = case_when(
      hour >= 0 & hour < 8  ~ "Night",
      hour >= 8 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon",
      hour >= 18 & hour < 24 ~ "Evening"
    ) # Time-Series Segmentation: Categorizing Time Into Four Periods
  ) %>% 
  select(uid, date, bin_dates, period, duration_new, type) # substitute hour by period

# test4 <- df_calllog_total %>% filter(uid == "REDACTED_ID_1@example")

# Perform multiple imputation
calllog_method <- make.method(df_filtered)
calllog_method[] <- ""
calllog_method["duration_new"] <- "pmm"

calllog_pred <- make.predictorMatrix(df_filtered)
calllog_pred[,] <- 0
calllog_pred["duration_new", c("bin_dates", "period", "type")] <- 1

mice_calllog <- mice(
  df_filtered,
  m = IMPUTATION_M,
  method = calllog_method,
  predictorMatrix = calllog_pred,
  remove.collinear = FALSE,
  seed = IMPUTATION_SEED
)

# Extract the first imputed dataset
completed_data <- complete(mice_calllog, 1)

# Calculate/Aggregate sum duration
df_calllog_total <- completed_data %>%
  group_by(uid, date) %>%
  summarize(
    total_call_duration = sum(duration_new, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  filter(uid %in% IDlist)

# Verify the results
verification <- df_calllog_total %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    n_hours = n() / n_distinct(date),
    .groups = "drop"
  )
table(verification$n_dates)

```

### Audio conversation

Total time spent on verbal conversations based on audio data to quantify in-person engagement.

```{r convo processing note}
# Note: Raw conversation processing is done in earlier scripts. We load the pre-processed conversation data below.
```

```{r fun expand_conversation_to_hours}
expand_conversation_to_hours <- function(start_time, end_time) {
  # Convert to POSIXct with explicit timezone
  start_time <- as.POSIXct(start_time, tz = "America/New_York")
  end_time <- as.POSIXct(end_time, tz = "America/New_York")
  
  # Calculate floor and ceiling times
  start_floor <- floor_date(start_time, "hour")
  end_ceiling <- ceiling_date(end_time, "hour")
  
  # Calculate hours as numeric differences to handle DST transition
  hours_seq <- seq(
    as.numeric(start_floor), 
    as.numeric(end_ceiling) - 3600, # subtract one hour in seconds
    by = 3600  # one hour in seconds
  )
  
  # Convert back to POSIXct
  as.POSIXct(hours_seq, origin = "1970-01-01", tz = "America/New_York")
}
```

```{r}
# Load processed df_convo_3
df_convo_3 <- read_csv(here::here("data", "df_convo_3_02202025.csv"))

# Test with one ID
# df_convo_3temp <- df_convo_3 %>% filter(uid %in% c("REDACTED_ID_1@example", "REDACTED_ID_2@example"))

# Step 1: Get clean date ranges from data_ema_6
date_ranges <- data_ema_6 %>%
  group_by(uid) %>%
  summarise(first_date = min(date), end_date = max(date), .groups = "drop")

# Step 2: Process conversation data with hours
df_convo_binned <- df_convo_3 %>%
  rowwise() %>%
  mutate(hours_spanned = list(expand_conversation_to_hours(start_time, end_time))) %>%
  unnest(hours_spanned) %>%
  mutate(
    date = as.Date(hours_spanned),
    hour = hour(hours_spanned)
  )

# Step 3: Create template with correct date ranges and bin_dates
template <- date_ranges %>%
  rowwise() %>%
  mutate(
    dates = list(seq(first_date, end_date, by = "day"))
  ) %>%
  unnest(dates) %>%
  group_by(uid) %>%
  mutate(
    bin_dates = row_number()
  ) %>%
  filter(bin_dates <= 90) %>%  # Keep only first 90 days
  crossing(hour = 0:23) %>%
  select(uid, date = dates, hour, bin_dates)

# Step 4: Join template with conversation data
df_convo_complete <- template %>%
  left_join(
    df_convo_binned,
    by = c("uid", "date", "hour")
  ) %>%
  group_by(uid) %>%
  mutate(
    convo_recorded = if_else(!is.na(convo_length_sec), 1, 0),
    x1 = if_else(is.na(x1), first(x1[!is.na(x1)]), x1),
    start_time = if_else(!is.na(convo_length_sec), start_time, NA),
    end_time = if_else(!is.na(convo_length_sec), end_time, NA),
    convo_length_sec = if_else(is.na(convo_length_sec), NA, convo_length_sec),
    convo_length_min = if_else(is.na(convo_length_min), NA, convo_length_min)
  ) %>%
  ungroup()

# Step 5: Combine with battery data
df_combined <- df_convo_complete %>%
  left_join(df_battery_complete, by = c("uid", "date", "hour")) %>%
  mutate(convo_recorded = ifelse(convo_recorded == 0 | is.na(convo_recorded), 0, 1)) 

# Step 6: Classify hours as before
df_combined_1 <- df_combined %>%
  mutate(
    non_event_status = case_when(
      convo_recorded == 0 & running == 0 ~ "Real Non-Event",
      convo_recorded == 0 & running == 1 ~ "Potential Missingness",
      convo_recorded == 1 & running == 0 ~ "Battery down with convo data",
      convo_recorded == 1 ~ "Conversation Recorded"
    )) 

table(df_combined_1$non_event_status)

# Step 7: Recode "Potential Missingness" to 0 sec as it's not detected IRL
df_convo_4 <- df_combined_1 %>%
  # filter(non_event_status == "Potential Missingness") %>%
  mutate(convo_length_sec = ifelse(non_event_status == "Potential Missingness", 0, convo_length_sec)) %>%
  select(uid, date, hour, bin_dates, start_time, end_time, 
         convo_length_sec, convo_length_min, convo_recorded, non_event_status) 

# Verification
verification <- df_convo_4 %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    n_hours = n() / n_distinct(date)
  )

table(verification$n_dates) # 300 pts and 90 days
```

#### Multiple imputation (Conversation)
```{r Multiple imputation - conversation}
# Transform to 4 periods
df_filtered <- df_convo_4 %>% 
  mutate(
    period = case_when(
      hour >= 0 & hour < 8  ~ "Night",
      hour >= 8 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon",
      hour >= 18 & hour < 24 ~ "Evening"
    ) # Time-Series Segmentation: Categorizing Time Into Four Periods
  ) %>% 
  select(uid, date, bin_dates, period, convo_length_sec, non_event_status) # keep bin_dates as time predictor

table(df_filtered$period, df_filtered$non_event_status)
sum(is.na(df_filtered$convo_length_sec)) # Only convo_length_sec has missingness

# Perform multiple imputation
convo_method <- make.method(df_filtered)
convo_method[] <- ""
convo_method["convo_length_sec"] <- "pmm"

convo_pred <- make.predictorMatrix(df_filtered)
convo_pred[,] <- 0
convo_pred["convo_length_sec", c("bin_dates", "period", "non_event_status")] <- 1

mice_convo <- mice(
  df_filtered,
  m = IMPUTATION_M, # Number of imputed datasets
  method = convo_method,
  predictorMatrix = convo_pred,
  remove.collinear = FALSE,
  seed = IMPUTATION_SEED
) # For reproducibility

# Extract the first imputed dataset
completed_data <- complete(mice_convo, 1)

sum(is.na(completed_data$convo_length_sec)) # Check if there's any NA

# Calculate/Aggregate sum duration
df_convo_total <- completed_data %>%
  group_by(uid, date) %>%
  summarize(
    total_convo_duration = sum(convo_length_sec, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  filter(uid %in% IDlist)

```

### App usage

```{r app categories from Anna}
# Read app categories data 
app_cat <- read_csv(here::here("data", "Apps_categorization_final.csv"))

# Keep any rows categorized as communication app in one of the categories
app_cat_1 <- app_cat %>% 
  filter(googleCats == "communication" | Schoedel == "Communication" | cats == "chat" | reducedCats == "Communication")

# Refine it to be more accurate
app_cat_2 <- app_cat %>% 
  filter(Schoedel == "Communication" | reducedCats == "Communication") %>% 
  filter(!cats %in% c("email", "videoconference", "other", "?")) %>% 
  select(ID = `...1`, everything())

# Re-categorize `com.viber.voip` to chat since it's a messaging and calling app
# The Viber messenger is very popular in the Philippines, Greece, Eastern Europe, Russia, the Middle East, and some Asian markets
app_cat_2$cats[app_cat_2$ID == 346] <- "chat"

# Remove all call & dialing in `cats` column since they're all calling / dialing log
app_cat_3 <- app_cat_2 %>% filter(!cats == "calling&contacts")

# Build a regex from escaped app IDs to avoid accidental partial matches
chat_apps_pattern <- paste(stringr::str_replace_all(app_cat_3$app_name, "([.|()\\^{}+$*?]|\\[|\\]|\\\\)", "\\\\\\1"), collapse = "|")

```

```{r app processing note}
# Note: Raw app data processing is done in earlier scripts.
```

```{r app initial clean}

# Load combined data
df_running_app <- read_csv(here::here("data", "df_running_app_combined.csv")) %>%
  select(1:4) 

# Transform to full datetime format
df_running_app$start_time <- as.POSIXct(df_running_app$day, format = "%Y-%m-%d %H:%M:%S", tz = APP_TIMEZONE)

# Filter the data frame for rows between 2021-06-01 and 2024-04-30
df_running_app <- df_running_app %>%
  filter(start_time >= as.Date("2021-06-01") & start_time <= as.Date("2024-04-30"))

# Capture social apps in chat_apps_pattern list
df_running_app$social_app <- ifelse(
  grepl(chat_apps_pattern, df_running_app$data),
  1, # Mark as social app
  0  # Not a social app
)

table(df_running_app$social_app)

# Final df
df_running_app_clean <- df_running_app %>% 
  select(event_id = x1, uid, app_data = data, start_time, social_app) %>% 
  filter(uid %in% IDlist)
```

```{r}
# Identify participants with no social app usage
no_social_app_users <- df_running_app_clean %>%
  group_by(uid) %>%
  summarize(all_social_app_zero = all(social_app == 0)) %>%
  filter(all_social_app_zero) %>%
  pull(uid)

cat("Participants with no social app usage: ")
print(no_social_app_users)
```

```{r}
# Step 1: Get clean date ranges from data_ema_6
date_ranges <- data_ema_6 %>%
  group_by(uid) %>%
  summarise(first_date = min(date), end_date = max(date), .groups = "drop")

# Step 2: Process app data with initial time binning
df_app_binned <- df_running_app_clean %>%
  mutate(
    start_time = as.POSIXct(start_time, tz = APP_TIMEZONE),
    date = as.Date(start_time),
    hour = hour(start_time),
    app_usage = 1
  )

# Step 3: Create template with correct date ranges and bin_dates
template <- date_ranges %>%
  rowwise() %>%
  mutate(
    dates = list(seq(first_date, end_date, by = "day"))
  ) %>%
  unnest(dates) %>%
  group_by(uid) %>%
  mutate(
    bin_dates = row_number()
  ) %>%
  filter(bin_dates <= 90) %>%  # Keep only first 90 days
  crossing(hour = 0:23) %>%
  select(uid, date = dates, hour, bin_dates)

# Step 4: Combine with battery data and app usage
df_app_complete <- template %>%
  left_join(df_battery_complete, by = c("uid", "date", "hour")) %>%
  left_join(df_app_binned, by = c("uid", "date", "hour")) %>%
  mutate(
    # Real non-event = battery off AND no app usage
    real_non_event = as.integer(running == 0 & is.na(app_usage)),
    # Fill NA app_usage with 0 where appropriate
    app_usage = ifelse(is.na(app_usage), 0, app_usage)
  )

table(df_app_complete$real_non_event)

# Step 4-1: create a df with only non-events
df_app_non_events <- df_app_complete %>% 
  filter(real_non_event == 1) %>% 
  select(uid, date, hour, bin_dates, real_non_event)

# Verification
verification <- df_app_complete %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    n_hours = n() / n_distinct(date),
    total_app_usage = sum(app_usage),
    n_real_non_events = sum(real_non_event)
  )

table(verification$n_dates)

# Check for any remaining NA in key columns
na_check <- df_app_complete %>%
  summarise(
    na_date = sum(is.na(date)),
    na_hour = sum(is.na(hour)),
    na_bin_dates = sum(is.na(bin_dates)),
    na_running = sum(is.na(running)),
    na_app_usage = sum(is.na(app_usage))
  )

print(na_check)
```


### Missingness imputation for app usage data
Note: Earlier app binning and non-event handling code has been consolidated into the sections above.

### Unlock data

```{r load unlock data, cache=FALSE}
# Load unlock data clean (raw processing done in earlier scripts)
load(here::here("results", "df_unlock_clean.RData"))
```

```{r}
df_unlock_clean <- df_unlock_clean %>% filter(uid %in% IDlist)
```

```{r}
# Some pts have less than 90 days of lock/unlock data
# Check the number of days per user
user_days <- df_unlock_clean %>%
  group_by(uid) %>%
  summarise(days = n_distinct(date))

# Find users with fewer than 90 days of data
missing_days_users <- user_days %>% filter(days < 90)
print(missing_days_users)
```

### Time Duration using Social App

#### Duration Calculation Algorithm
```{r}
# Filter for only social app events
# social_app_events <- df_running_app_clean %>%
#   #filter(social_app == 1) %>%
#   arrange(uid, start_time)

# Filter for only social app events using binned df
social_app_events <- df_app_complete %>%
  filter(social_app == 1) %>%
  arrange(uid, start_time)

# social_non_event <- social_app_events %>% filter(social_app_events$real_non_event == 1)
# 
# # Create datetime
# social_non_event_1 <- social_non_event %>%
#   mutate(
#     day = as.POSIXct(
#       paste(date, sprintf("%02d:00:00.00", hour)),
#       format = "%Y-%m-%d %H:%M:%OS",
#       tz = "America/New_York"  # or your preferred timezone
#     )
#   ) %>% 
#   select(1:4, 13)


# Define the timeout duration (5 minutes)
timeout_seconds <- SOCIAL_APP_TIMEOUT_SEC

# Function to calculate durations for each social app event
calculate_durations <- function(event_start_time, unlock_data, next_event_time) {
  # Convert unlock_data to a data frame of seconds and lock/unlock status
  unlock_df <- data.frame(
    timestamp = seq(
    as.POSIXct(paste(as.Date(event_start_time), "00:00:00"), tz = APP_TIMEZONE),
    by = "1 sec",
    length.out = length(unlock_data)
  ),
    is_unlocked = unlock_data
  )
  
  # Filter data for relevant seconds
  relevant_data <- unlock_df %>% filter(timestamp >= floor_date(event_start_time, "second")) # rounding event_start_time to the nearest second
  
  # Condition 1: Any new app opened
  if (length(next_event_time) == 0 || is.na(next_event_time)) {
    condition1_duration <- Inf
  } else {
    condition1_duration <- as.numeric(difftime(next_event_time, event_start_time, units = "secs"))
  }
  
  # Condition 2: Phone locked
  condition2_time <- relevant_data %>%
    filter(is_unlocked == 0) %>% # lock: 0, unlock: 1
    slice(1) %>%
    pull(timestamp)
  
  # Handle case where no phone lock occurs
  condition2_duration <- if (length(condition2_time) == 0) {
    Inf # use Inf as we will find the minimum value later
  } else {
    as.numeric(difftime(condition2_time, event_start_time, units = "secs"))
  }
  
  # Condition 3: Timeout (5 minutes)
  condition3_duration <- timeout_seconds
  
  # Calculate the minimum duration
  min_duration <- min(condition1_duration, condition2_duration, condition3_duration)
  
  return(data.frame(
    condition1_duration = condition1_duration,
    condition2_duration = condition2_duration,
    condition3_duration = condition3_duration,
    min_duration = min_duration
  ))
}
```

#### Test in a small batch of pts
```{r social app + text - small batch of pts}
# small dataset to test
test_id <- IDlist[1:2]
# Process each user's social app events
results <- list()

# Loop through each unique user ID
for (id in test_id) {
  
  # Filter data for this user
  user_social_events <- social_app_events %>% filter(uid == id)
  user_all_events <- df_running_app_clean %>% filter(uid == id) %>% arrange(start_time)
  user_unlock <- df_unlock_clean %>% filter(uid == id)
  
  # Skip if no social events are available for the user
  if (nrow(user_social_events) == 0) {
    next
  }
  
  # Skip if no unlock data is available for the user
  if (nrow(user_unlock) == 0) {
    next
  }
  
  # Loop through each social app event for this user
  for (i in 1:nrow(user_social_events)) {
    
    # Get social app event details
    event_start_time <- user_social_events$start_time[i]
    
    # Extract the date from the event_start_time
    event_date <- as.Date(event_start_time)
    
    # Skip if event_date is NA or not in user_unlock$date
    if (is.na(event_date) || nrow(user_unlock) == 0 || !event_date %in% user_unlock$date) {
      next
    }
    
    # Find the unlock data corresponding to the event's date
    # if (!event_date %in% user_unlock$date) {
    #   # Skip if the event_date is not in the user's unlock data
    #   next
    # }
    
    # Skip if the event_date is not in user_unlock$date
    # if (!event_date %in% user_unlock$date) {
    #   next
    # }
    
    # Extract and unlist the unlock data for the event's date
    unlock_data <- user_unlock %>%
      filter(date == event_date) %>%
      pull(data) %>%
      unlist()
    
    # Skip if no unlock data is available for the date
    if (length(unlock_data) == 0) {
      next
    }
    
    # Get the next event start time (ANY app, including non-social apps)
    next_event_time <- user_all_events %>%
      filter(start_time > event_start_time) %>%
      slice(1) %>%
      pull(start_time)
    
    # Handle case where there is no next event
    if (length(next_event_time) == 0) {
      next_event_time <- NA
    }
    
    # Calculate durations for this event
    durations <- calculate_durations(event_start_time, unlock_data, next_event_time)
    durations$uid <- id
    durations$event_id <- user_social_events$event_id[i]
    durations$date <- event_date               # Add the date column
    durations$event_start_time <- event_start_time  # Add event_start_time column
    
    # Append to results
    results[[length(results) + 1]] <- durations
  }
}

# Combine all results from a list of data frames into a single data frame
results_df1 <- do.call(rbind, results)

# Reorder df
results_df2 <- results_df1 %>% select(uid, event_id, date, event_start_time, everything())

# Calculate total duration each day per participant
results_df3 <- results_df2 %>% 
  group_by(uid, date) %>% 
  summarise(total_duration = sum(min_duration)) # If `total_duration` is negative value, it should equal to 0 (all the negative value is smaller than 1 sec)
```

#### All pts

```{r social app + text - all pts}
# Process each user's social app events
results <- list()

# Loop through each unique user ID
for (id in IDlist) {
  
  # Filter data for this user
  user_social_events <- social_app_events %>% filter(uid == id)
  user_all_events <- df_running_app_clean %>% filter(uid == id) %>% arrange(start_time)
  user_unlock <- df_unlock_clean %>% filter(uid == id)
  
  # Skip if no social events are available for the user
  if (nrow(user_social_events) == 0) {
    next
  }
  
  # Skip if no unlock data is available for the user
  if (nrow(user_unlock) == 0) {
    next
  }
  
  # Loop through each social app event for this user
  for (i in 1:nrow(user_social_events)) {
    
    real_non_event <- user_social_events$real_non_event[i]
    # Get social app event details
    event_start_time <- user_social_events$start_time[i]
    
    # Extract the date from the event_start_time
    event_date <- as.Date(event_start_time)
    
    # Skip if event_date is NA or not in user_unlock$date
    if (is.na(event_date) || nrow(user_unlock) == 0 || !event_date %in% user_unlock$date) {
      next
    }
    
    # Extract and unlist the unlock data for the event's date
    unlock_data <- user_unlock %>%
      filter(date == event_date) %>%
      pull(data) %>%
      unlist()

    # Skip if no unlock data is available for the date
    if (length(unlock_data) == 0) {
      next
    }

    # Get the next event start time (ANY app, including non-social apps)
    next_event_time <- user_all_events %>%
      filter(start_time > event_start_time) %>%
      slice(1) %>%
      pull(start_time)

    # Handle case where there is no next event
    if (length(next_event_time) == 0) {
      next_event_time <- NA
    }

    # Calculate durations for this event
    durations <- calculate_durations(event_start_time, unlock_data, next_event_time)
    durations$uid <- id
    durations$event_id <- user_social_events$event_id[i]
    durations$date <- event_date               # Add the date column
    durations$event_start_time <- event_start_time  # Add event_start_time column

    # Append to results
    results[[length(results) + 1]] <- durations
  }
}

# Combine all results into a single data frame
results_all <- do.call(rbind, results)

# Reorder data frame columns
results_all <- results_all %>%
  select(uid, event_id, date, event_start_time, everything())

```


```{r}
results_all <- read_csv(here::here("data", "result_app_duration_all.csv"))

results_all_1 <- results_all %>% 
  mutate(hour = hour(event_start_time)) %>% 
  select(uid, date, hour, condition1_duration, condition2_duration, condition3_duration)
 
df_app_non_events_1 <- df_app_non_events %>% 
  mutate(
    condition1_duration = NA, 
    condition2_duration = NA, 
    condition3_duration = 300
  ) %>% 
  select(-bin_dates, -real_non_event)

results_all_2 <- rbind(results_all_1, df_app_non_events_1)

# Check for duplicates specifically when both duration conditions are NA:
results_all_2 %>% 
  filter(is.na(condition1_duration) & is.na(condition2_duration)) %>%
  group_by(uid, date, hour) %>%
  summarise(n = n(), .groups = "drop") %>%
  filter(n > 1)



# results_all_2 <- results_all_1 %>% 
#   left_join(social_non_event_1, by = c("uid", "date", "hour")) %>% 
#   select(1:10, 19) %>% 
#   arrange(uid, date, hour)


```

```{r}
df_filtered_nonevent <- df_app_non_events_1 %>% 
  # filter(type %in% c("1", "2", "7")) %>% 
  mutate(
    period = case_when(
      hour >= 0 & hour < 8  ~ "Night",
      hour >= 8 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon",
      hour >= 18 & hour < 24 ~ "Evening"
    ) # Time-Series Segmentation: Categorizing Time Into Four Periods
  ) %>% 
  select(uid, date, hour, period, condition1_duration, condition2_duration, condition3_duration) # substitute hour by period

```

```{r}

df_filtered <- results_all_2 %>% 
  # filter(type %in% c("1", "2", "7")) %>% 
  mutate(
    period = case_when(
      hour >= 0 & hour < 8  ~ "Night",
      hour >= 8 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon",
      hour >= 18 & hour < 24 ~ "Evening"
    ) # Time-Series Segmentation: Categorizing Time Into Four Periods
  ) %>% 
  mutate(
    condition1_duration = ifelse(is.infinite(condition1_duration), NA, 
                                 ifelse(condition1_duration < 0, 0, condition1_duration)),
    condition2_duration = ifelse(is.infinite(condition2_duration), NA, 
                                 ifelse(condition2_duration < 0, 0, condition2_duration)),
    condition3_duration = ifelse(is.infinite(condition3_duration), NA, 
                                 ifelse(condition3_duration < 0, 0, condition3_duration))
  ) %>% 
  select(uid, date, hour, period, condition1_duration, condition2_duration, condition3_duration) # substitute hour by period

# Check initial data
print("Initial NA counts:")
print(colSums(is.na(df_filtered)))

# Check for Inf and negative values
colSums(sapply(df_filtered[, c("condition1_duration", "condition2_duration", "condition3_duration")], is.infinite))

# Run mice() on all three duration candidates with explicit predictors
app_method <- make.method(df_filtered)
app_method[] <- ""
app_method[c("condition1_duration", "condition2_duration", "condition3_duration")] <- "pmm"

app_pred <- make.predictorMatrix(df_filtered)
app_pred[,] <- 0
app_pred[c("condition1_duration", "condition2_duration", "condition3_duration"),
         c("hour", "period", "condition1_duration", "condition2_duration", "condition3_duration")] <- 1
diag(app_pred) <- 0

mice_app <- mice(
  df_filtered,
  method = app_method,
  predictorMatrix = app_pred,
  remove.collinear = FALSE,
  m = IMPUTATION_M,
  maxit = 50,
  seed = IMPUTATION_SEED
)

# Extract the completed data
completed_data <- complete(mice_app, 1)

# Check final data
print("Final NA counts:")
print(colSums(is.na(completed_data)))

```
```{r}
# Calculate/Aggregate sum duration
df_app_min <- completed_data %>%
  mutate(min_duration = pmin(condition1_duration, condition2_duration, condition3_duration, na.rm = T)) 

df_app_total <- df_app_min %>% 
  group_by(uid, date) %>%
  summarise(total_duration = sum(min_duration, na.rm = TRUE)) %>% 
  filter(uid %in% IDlist)

```

```{r}
# Step 1: Get date ranges from data_ema_6
date_ranges <- data_ema_6 %>%
  group_by(uid) %>%
  summarise(first_date = min(date), end_date = max(date), .groups = "drop")

# Step 2: Create template with correct date ranges and bin_dates
template <- date_ranges %>%
  rowwise() %>%
  mutate(
    dates = list(seq(first_date, end_date, by = "day"))
  ) %>%
  unnest(dates) %>%
  group_by(uid) %>%
  mutate(
    bin_dates = row_number()
  ) %>%
  filter(bin_dates <= 90) %>%  # Keep only first 90 days
  select(uid, date = dates, bin_dates)

# Step 3: Join with app data
df_app_total_complete <- template %>%
  left_join(df_app_total, by = c("uid", "date")) %>%
  mutate(
    total_duration = if_else(is.na(total_duration), 0, total_duration)  # Replace NA with 0
  )

# Check if any NA
print(colSums(is.na(df_app_total_complete)))

# Verification
verification <- df_app_total_complete %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    min_date = min(date),
    max_date = max(date),
    n_bins = n_distinct(bin_dates)
  )

table(verification$n_dates)
print(head(verification))
```

### Combine data of total duration to a df
```{r}

# df_app_total <- read_csv("./data/df_app_total.csv") 

df_app_total <- df_app_total %>%
  mutate(total_duration = ifelse(total_duration < 0, 0, total_duration),
         total_duration = round(total_duration))
  
  
# Merge the three data frames based on `uid` and `date`
df_combined_total <- data_ema_6 %>%
  left_join(df_app_total_complete %>% select(-bin_dates), by = c("uid", "date")) %>%
  left_join(df_calllog_total, by = c("uid", "date")) %>%
  left_join(df_convo_total, by = c("uid", "date")) %>%
  select(uid, date, bin_dates, daily_phq_total, total_call_duration, total_app_duration = total_duration, total_convo_duration) %>%
  mutate(daily_phq_total = round(daily_phq_total),
    total_app_duration = round(total_app_duration)) # round to integer

# Check NA
print(colSums(is.na(df_combined_total))) # 27000 rows

# Verification
verification <- df_combined_total %>%
  group_by(uid) %>%
  summarise(
    n_dates = n_distinct(date),
    min_date = min(date),
    max_date = max(date),
    n_bins = n_distinct(bin_dates)
  )

table(verification$n_dates)
print(head(verification))

```

# Calculation

```{r}
final_ratio <- df_combined_total %>%
  group_by(uid) %>%
  summarise(
    total_convo = sum(total_convo_duration, na.rm = TRUE),
    total_other = sum(total_call_duration + total_app_duration, na.rm = TRUE),
    convo_ratio = total_convo / (total_convo + total_other)
  ) %>%
  ungroup()

# Add some descriptive statistics
summary_stats <- final_ratio %>%
  summarise(
    mean_ratio = mean(convo_ratio, na.rm = TRUE),
    median_ratio = median(convo_ratio, na.rm = TRUE),
    sd_ratio = sd(convo_ratio, na.rm = TRUE),
    min_ratio = min(convo_ratio, na.rm = TRUE),
    max_ratio = max(convo_ratio, na.rm = TRUE)
  )

# Print results
print("Summary of conversation ratio per participant:")
print(summary_stats)

# Optional: Look at distribution
hist_data <- ggplot(final_ratio, aes(x = convo_ratio)) +
  geom_histogram(binwidth = 0.05) +
  labs(
    title = "Distribution of Verbal Conversation Ratio",
    x = "Ratio of Verbal Conversation Time to Total Communication Time",
    y = "Count"
  )

print(hist_data)
```

```{r}

library(wesanderson)

df_combined_total <- df_combined_total %>%
  mutate(social_interaction_ratio = total_convo_duration / 
         (total_convo_duration + total_call_duration + total_app_duration))

# Compute the average ratio & PHQ score per day
df_avg_ratio <- df_combined_total %>%
  group_by(bin_dates) %>%
  summarize(
    mean_ratio = mean(social_interaction_ratio, na.rm = TRUE),
    mean_phq = mean(daily_phq_total, na.rm = TRUE)
  )

# Plot both lines
ggplot(df_avg_ratio, aes(x = bin_dates)) +
  geom_line(aes(y = mean_ratio, color = "Social Interaction Ratio"), size = 1) +
  geom_line(aes(y = mean_phq / max(mean_phq, na.rm = TRUE), color = "Daily PHQ Score"), size = 1) + # , linetype = "dashed"
  scale_color_manual(values = c("Social Interaction Ratio" = "#E6A8B5", "Daily PHQ Score" = "#8BAF9F")) +
  labs(title = "Social Interaction Ratio & Depression Score Over 90 Days",
       x = "Days",
       y = "Scaled Values (PHQ Score Scaled)",
       color = "Legend") +
  theme_minimal()

```

```{r}
# Plot 1: Box plot by time series to see temporal patterns
time_series_plot <- df_combined_total %>%
  mutate(
    daily_ratio = total_convo_duration / (total_convo_duration + total_call_duration + total_app_duration),
    week = ceiling(bin_dates/7)  # Group by weeks
  ) %>%
  ggplot(aes(x = factor(week), y = daily_ratio)) +
  geom_boxplot(outlier.alpha = 0.5) +
  labs(
    title = "Distribution of Daily Conversation Ratio by Week",
    x = "Week Number",
    y = "Ratio of Verbal to Total Communication",
    caption = "Box plots show median, quartiles, and outliers for each week"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0))

# Plot 2: Stacked area chart to show relative proportions over time
composition_plot <- df_combined_total %>%
  group_by(bin_dates) %>%
  summarise(
    verbal = mean(total_convo_duration, na.rm = TRUE),
    calls = mean(total_call_duration, na.rm = TRUE),
    apps = mean(total_app_duration, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(verbal, calls, apps), 
               names_to = "type", 
               values_to = "duration") %>%
  ggplot(aes(x = bin_dates, y = duration, fill = type)) +
  geom_area(position = "fill") +
  scale_fill_brewer(palette = "Blues", 
                    labels = c("App Usage", "Call Duration", "Verbal Conversation")) +
  labs(
    title = "Daily Composition of Communication Types",
    x = "Day Number",
    y = "Proportion of Total Communication Time",
    fill = "Communication Type"
  ) +
  theme_minimal()

# Display plots
print(time_series_plot)
print(composition_plot)
```

```{r}
# 1. Individual Trajectory Plot with Mean Trend
trajectory_plot <- df_combined_total %>%
  mutate(
    daily_ratio = total_convo_duration / (total_convo_duration + total_call_duration + total_app_duration)
  ) %>%
  ggplot(aes(x = bin_dates)) +
  # Add individual trajectories with low alpha
  geom_line(aes(y = daily_ratio, group = uid), alpha = 0.1, color = "gray50") +
  # Add mean trend line
  stat_summary(aes(y = daily_ratio), fun = mean, geom = "line", 
              color = "blue", size = 1.5) +
  # Add confidence interval
  stat_summary(aes(y = daily_ratio), fun.data = mean_cl_normal, 
              geom = "ribbon", alpha = 0.2) +
  labs(
    title = "Individual Trajectories with Population Mean",
    x = "Day Number",
    y = "Ratio of Verbal to Total Communication",
    caption = "Gray lines: individual participants; Blue line: population mean"
  ) +
  theme_minimal()

# 2. Heatmap of Daily Patterns
heatmap_plot <- df_combined_total %>%
  mutate(
    daily_ratio = total_convo_duration / (total_convo_duration + total_call_duration + total_app_duration),
    week = ceiling(bin_dates/7),  # Group into weeks
    weekday = wday(date, label = TRUE)  # Get day of week
  ) %>%
  group_by(week, weekday) %>%
  summarise(
    mean_ratio = mean(daily_ratio, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  ggplot(aes(x = weekday, y = factor(week), fill = mean_ratio)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    title = "Weekly Patterns of Communication Ratio",
    x = "Day of Week",
    y = "Week Number",
    fill = "Mean Ratio"
  ) +
  theme_minimal()

# 3. Moving Average Trend with Quantiles
trend_plot <- df_combined_total %>%
  mutate(
    daily_ratio = total_convo_duration / (total_convo_duration + total_call_duration + total_app_duration)
  ) %>%
  group_by(bin_dates) %>%
  summarise(
    median = median(daily_ratio, na.rm = TRUE),
    q25 = quantile(daily_ratio, 0.25, na.rm = TRUE),
    q75 = quantile(daily_ratio, 0.75, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = bin_dates)) +
  geom_line(aes(y = median), color = "blue") +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.2) +
  labs(
    title = "Temporal Trend of Communication Ratio",
    x = "Day Number",
    y = "Ratio of Verbal to Total Communication",
    caption = "Blue line: median; Shaded area: 25th-75th percentile range"
  ) +
  theme_minimal()

# Print plots
print(trajectory_plot)
print(heatmap_plot)
print(trend_plot)
```

```{r}
library(ggplot2)
library(dplyr)

df_combined_total <- df_combined_total %>%
  mutate(social_interaction_ratio = total_convo_duration / 
         (total_convo_duration + total_call_duration + total_app_duration))

# Compute the average ratio & PHQ score per day
df_avg_ratio <- df_combined_total %>%
  group_by(bin_dates) %>%
  summarize(
    mean_ratio = mean(social_interaction_ratio, na.rm = TRUE),
    mean_phq = mean(daily_phq_total, na.rm = TRUE)
  )

# Plot both lines
ggplot(df_avg_ratio, aes(x = bin_dates)) +
  geom_line(aes(y = mean_ratio, color = "Social Interaction Ratio"), size = 1) +
  geom_line(aes(y = mean_phq / max(mean_phq, na.rm = TRUE), color = "Daily PHQ Score"), size = 1) + # , linetype = "dashed"
  scale_color_manual(values = c("Social Interaction Ratio" = "blue", "Daily PHQ Score" = "red")) +
  labs(title = "Social Interaction Ratio & Depression Score Over 90 Days",
       x = "Days",
       y = "Scaled Values (PHQ Score Scaled)",
       color = "Legend") +
  theme_minimal()

```


# Export All Imputed Datasets for Rubin's Rules Pooling

This section creates m=5 versions of the combined dataset, one per imputation.
The analysis script will run mlVAR on each and pool results using Rubin's rules.

```{r create-pooled-datasets}
# ---- Helper functions to process each imputation ----

process_ema_imp <- function(mice_obj, imp_idx) {
  data_ema_3 <- complete(mice_obj, imp_idx)
  data_ema_4 <- data_ema_3 %>% filter(bin_dates %in% 1:STUDY_DAYS)
  data_ema_5 <- data_ema_4 %>%
    mutate(phq_total = rowSums(select(., paste0("phq_", 1:9)), na.rm = TRUE),
           phq_mean = rowMeans(select(., paste0("phq_", 1:9)), na.rm = TRUE))
  data_ema_6 <- data_ema_5 %>%
    group_by(uid, date) %>%
    summarize(
      daily_phq_total = mean(phq_total, na.rm = TRUE),
      daily_phq_mean = mean(phq_mean, na.rm = TRUE),
      bin_dates = first(bin_dates),
      .groups = "drop"
    )
  return(data_ema_6)
}

process_calllog_imp <- function(mice_obj, imp_idx, id_list) {
  completed <- complete(mice_obj, imp_idx)
  df_total <- completed %>%
    group_by(uid, date) %>%
    summarize(
      total_call_duration = sum(duration_new, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(uid %in% id_list)
  return(df_total)
}

process_convo_imp <- function(mice_obj, imp_idx, id_list) {
  completed <- complete(mice_obj, imp_idx)
  df_total <- completed %>%
    group_by(uid, date) %>%
    summarize(
      total_convo_duration = sum(convo_length_sec, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    filter(uid %in% id_list)
  return(df_total)
}

process_app_imp <- function(mice_obj, imp_idx, id_list) {
  completed <- complete(mice_obj, imp_idx)
  df_app_min <- completed %>%
    mutate(min_duration = pmin(condition1_duration, condition2_duration,
                               condition3_duration, na.rm = TRUE))
  df_total <- df_app_min %>%
    group_by(uid, date) %>%
    summarise(total_duration = sum(min_duration, na.rm = TRUE), .groups = "drop") %>%
    filter(uid %in% id_list)
  return(df_total)
}

# ---- Generate all m combined datasets ----

# Date ranges and template (same across imputations since dates are not imputed)
date_ranges_pool <- data_ema_6 %>%
  group_by(uid) %>%
  summarise(first_date = min(date), end_date = max(date), .groups = "drop")

template_pool <- date_ranges_pool %>%
  rowwise() %>%
  mutate(dates = list(seq(first_date, end_date, by = "day"))) %>%
  unnest(dates) %>%
  group_by(uid) %>%
  mutate(bin_dates = row_number()) %>%
  filter(bin_dates <= STUDY_DAYS) %>%
  select(uid, date = dates, bin_dates) %>%
  ungroup()

df_combined_all <- map_dfr(1:IMPUTATION_M, function(i) {
  cat("Processing imputation", i, "of", IMPUTATION_M, "\n")

  ema_i <- process_ema_imp(mice_ema, i)
  calllog_i <- process_calllog_imp(mice_calllog, i, IDlist)
  convo_i <- process_convo_imp(mice_convo, i, IDlist)
  app_i <- process_app_imp(mice_app, i, IDlist)

  # Complete app data with template (fill missing dates with 0)
  app_complete_i <- template_pool %>%
    left_join(app_i, by = c("uid", "date")) %>%
    mutate(total_duration = if_else(is.na(total_duration), 0, total_duration))

  # Clean app durations
  app_complete_i <- app_complete_i %>%
    mutate(total_duration = ifelse(total_duration < 0, 0, total_duration),
           total_duration = round(total_duration))

  # Combine all data sources (drop bin_dates from app to avoid suffix conflict)
  combined_i <- ema_i %>%
    left_join(app_complete_i %>% select(-bin_dates), by = c("uid", "date")) %>%
    left_join(calllog_i, by = c("uid", "date")) %>%
    left_join(convo_i, by = c("uid", "date")) %>%
    select(uid, date, bin_dates,
           daily_phq_total, daily_phq_mean,
           total_call_duration,
           total_app_duration = total_duration,
           total_convo_duration) %>%
    mutate(daily_phq_total = round(daily_phq_total),
           total_app_duration = round(total_app_duration),
           .imp = i)

  return(combined_i)
})

cat("Total rows:", nrow(df_combined_all), "\n")
cat("Expected:", 300 * STUDY_DAYS * IMPUTATION_M, "\n")
cat("Imputations:", unique(df_combined_all$.imp), "\n")
```

```{r export-pooled-data}
# Export stacked dataset for Rubin's rules analysis
write_csv(df_combined_all, here::here("data", "df_combined_total_pooled.csv"))
cat("Exported pooled dataset with", IMPUTATION_M, "imputations to data/df_combined_total_pooled.csv\n")

# Also keep the single-imputation export for backwards compatibility
write_csv(
  df_combined_all %>% filter(.imp == 1) %>% select(-.imp),
  here::here("data", "df_combined_total.csv")
)
cat("Exported single-imputation dataset to data/df_combined_total.csv\n")
```
